# Benthic Invertebrate Indices {#benthos_index}

```{r echo=FALSE}
knitr::opts_chunk$set(echo = F)
library(ecodata)
library(downloadthis)
```

**Last updated**: April 08, 2026

**Description**: Aggregate macrobenthos and megabenthos invertebrate indices from fish stomach contents

**Indicator family**: 

- [X] Lower trophic levels


**Contributor(s)**: Maxwell Grezlik, Sarah Gaichas, James Gartland, Brian E. Smith, Sarah Weisberg, Sean Lucey

**Affiliations**: NEFSC

```{r download_benthos_index}
# Download dataset from ecodata
downloadthis::download_this(
ecodata::benthos_index,
output_name = 'benthos_index',
output_extension = '.csv',
button_label = 'Download dataset from `ecodata`',
button_type = 'default',
has_icon = TRUE,
icon = 'fa fa-save',
class = 'hvr-sweep-to-left',
csv2 = F)
```

## Introduction to Indicator
Benthic invertebrates provide supporting ecosystem services as important prey for many economically important fish species on the Northeast US shelf.  Changes to the benthic invertebrate forage base could have important implications for regional food webs. However, we lack direct measurements of benthic invertebrate biomass for most unexploited taxa. Spatially explicit indices of abundance for benthos groups were estimated using spatio-temporal modeling (VAST, @thorson_comparing_2017; @thorson_guidance_2019).

We define Macrobenthos as bottom-dwelling invertebrates retained on a 0.5-1.0 mm sieve. These include polychaete worms, small crustaceans, bivalves (non-commercial), gastropods, nemerteans, tunicates, cnidarians,  brittle stars, sea cucumbers, and sand dollars.

We define Megabenthos as sea stars (Echinodermata: Asteroidea) and large Arthropods (horseshoe crabs (Merostomata), mantis shrimp (Crustacea: Stomatopoda), and crabs (Crustacea: Decapoda: Brachyura and Anomura other than hermit crabs)). 

The indices developed here use similar methods to the [Forage Fish Index](https://noaa-edab.github.io/catalog/forage_index.html). The spatially-explicit macrobenthos and megabenthos indices estimate the combined biomass of 833 and 105 taxa, respectively, using stomach contents information from 88 predator/size combinations from fish collected on bottom trawl surveys.

In addition to an index of benthic invertebrate biomass in each EPU, the coastwide center of gravity for the combined benthic biomass was estimated to evaluate whether distributions of benthic invertebrate biomass have changed over time.

## Key Results and Visualizations
Time series extend from 1980-2023.  Macrobenthos and Megabenthos are included on separate plots for each EPU.

Variables plotted are seasonal (Fall or Spring) Macrobenthos or Megabenthos Biomass Estimates with standard error bands.

Eastward and Northward components of the Center of Gravity, along with the Center of Gravity SE, are plotted for Fall and Spring Macrobenthos or Megabenthos Biomass.

**Mid-Atlantic, Megabenthos, index**
```{r plot_benthos_indexMidAtlanticMegabenthosindex}
# Plot indicator
ggplotObject <- ecodata::plot_benthos_index(report = 'MidAtlantic', varName = 'Megabenthos', plottype = 'index', n = 10)
ggplotObject
try(downloadthis::download_this(
ggplotObject$data,
output_name = 'benthos_indexMidAtlanticMegabenthosindex',
output_extension = '.csv',
button_label = 'Download the data plotted above',
button_type = 'default',
has_icon = TRUE,
icon = 'fa fa-save',
class = 'hvr-sweep-to-left',
csv2 = F),
silent = TRUE)
```

**New England, Megabenthos, index**
```{r plot_benthos_indexNewEnglandMegabenthosindex}
# Plot indicator
ggplotObject <- ecodata::plot_benthos_index(report = 'NewEngland', varName = 'Megabenthos', plottype = 'index', n = 10)
ggplotObject
try(downloadthis::download_this(
ggplotObject$data,
output_name = 'benthos_indexNewEnglandMegabenthosindex',
output_extension = '.csv',
button_label = 'Download the data plotted above',
button_type = 'default',
has_icon = TRUE,
icon = 'fa fa-save',
class = 'hvr-sweep-to-left',
csv2 = F),
silent = TRUE)
```

**Mid-Atlantic, Macrobenthos, index**
```{r plot_benthos_indexMidAtlanticMacrobenthosindex}
# Plot indicator
ggplotObject <- ecodata::plot_benthos_index(report = 'MidAtlantic', varName = 'Macrobenthos', plottype = 'index', n = 10)
ggplotObject
try(downloadthis::download_this(
ggplotObject$data,
output_name = 'benthos_indexMidAtlanticMacrobenthosindex',
output_extension = '.csv',
button_label = 'Download the data plotted above',
button_type = 'default',
has_icon = TRUE,
icon = 'fa fa-save',
class = 'hvr-sweep-to-left',
csv2 = F),
silent = TRUE)
```

**New England, Macrobenthos, index**
```{r plot_benthos_indexNewEnglandMacrobenthosindex}
# Plot indicator
ggplotObject <- ecodata::plot_benthos_index(report = 'NewEngland', varName = 'Macrobenthos', plottype = 'index', n = 10)
ggplotObject
try(downloadthis::download_this(
ggplotObject$data,
output_name = 'benthos_indexNewEnglandMacrobenthosindex',
output_extension = '.csv',
button_label = 'Download the data plotted above',
button_type = 'default',
has_icon = TRUE,
icon = 'fa fa-save',
class = 'hvr-sweep-to-left',
csv2 = F),
silent = TRUE)
```

**Mid-Atlantic, Megabenthos, cog**
```{r plot_benthos_indexMidAtlanticMegabenthoscog}
# Plot indicator
ggplotObject <- ecodata::plot_benthos_index(report = 'MidAtlantic', varName = 'Megabenthos', plottype = 'cog', n = 10)
ggplotObject
try(downloadthis::download_this(
ggplotObject$data,
output_name = 'benthos_indexMidAtlanticMegabenthoscog',
output_extension = '.csv',
button_label = 'Download the data plotted above',
button_type = 'default',
has_icon = TRUE,
icon = 'fa fa-save',
class = 'hvr-sweep-to-left',
csv2 = F),
silent = TRUE)
```

**New England, Megabenthos, cog**
```{r plot_benthos_indexNewEnglandMegabenthoscog}
# Plot indicator
ggplotObject <- ecodata::plot_benthos_index(report = 'NewEngland', varName = 'Megabenthos', plottype = 'cog', n = 10)
ggplotObject
try(downloadthis::download_this(
ggplotObject$data,
output_name = 'benthos_indexNewEnglandMegabenthoscog',
output_extension = '.csv',
button_label = 'Download the data plotted above',
button_type = 'default',
has_icon = TRUE,
icon = 'fa fa-save',
class = 'hvr-sweep-to-left',
csv2 = F),
silent = TRUE)
```

**Mid-Atlantic, Macrobenthos, cog**
```{r plot_benthos_indexMidAtlanticMacrobenthoscog}
# Plot indicator
ggplotObject <- ecodata::plot_benthos_index(report = 'MidAtlantic', varName = 'Macrobenthos', plottype = 'cog', n = 10)
ggplotObject
try(downloadthis::download_this(
ggplotObject$data,
output_name = 'benthos_indexMidAtlanticMacrobenthoscog',
output_extension = '.csv',
button_label = 'Download the data plotted above',
button_type = 'default',
has_icon = TRUE,
icon = 'fa fa-save',
class = 'hvr-sweep-to-left',
csv2 = F),
silent = TRUE)
```

**New England, Macrobenthos, cog**
```{r plot_benthos_indexNewEnglandMacrobenthoscog}
# Plot indicator
ggplotObject <- ecodata::plot_benthos_index(report = 'NewEngland', varName = 'Macrobenthos', plottype = 'cog', n = 10)
ggplotObject
try(downloadthis::download_this(
ggplotObject$data,
output_name = 'benthos_indexNewEnglandMacrobenthoscog',
output_extension = '.csv',
button_label = 'Download the data plotted above',
button_type = 'default',
has_icon = TRUE,
icon = 'fa fa-save',
class = 'hvr-sweep-to-left',
csv2 = F),
silent = TRUE)
```


## Indicator statistics 
Spatial scale: by EPU and full shelf

Temporal scale: Spring (January-June), Fall (July-December)

**Synthesis Theme**:

- [X] Multiple System Drivers


```{r autostats_benthos_index}
# Either from Contributor or ecodata
```

## Implications
Macrobenthos indices show long term declines in spring in all EPUs. In contrast, Megabenthos indices show long term increases in fall in GOM and GB, and in spring in MAB. Megabenthos indices also show a recent trend in spring in GB.

Center of gravity indicators show recent trends towards the north and east in spring for Megabenthos in all EPUs. Macrobenthos center of gravity shows a significant long term trend towards the west over time. This is the opposite direction observed for aggregate forage fish and surveyed fish species in aggregate.

## Get the data

**Point of contact**: [EDAB, edab.data@noaa.gov](mailto:EDAB, edab.data@noaa.gov){.email}

**ecodata name**: `ecodata::benthos_index`

**Variable definitions**

1. Spring Macrobenthos Biomass Estimate = aggregate macrobenthos biomass months 1-6, units relative grams per stomach 2. Spring Macrobenthos Biomass Estimate SE = standard error of aggregate macrobenthos biomass months 1-6, units relative grams per stomach 
3. Fall Macrobenthos Biomass Estimate = aggregate macrobenthos biomass months 7-12, units relative grams per stomach 4. Fall Macrobenthos Biomass Estimate SE = standard error of aggregate macrobenthos biomass months 7-12, units relative grams per stomach 
5. Fall Eastward Macrobenthos Center of Gravity = average eastward location of macrobenthos biomass months 7-12, units kilometers 6. Fall Eastward Macrobenthos Center of Gravity SE = standard error of average eastward location of macrobenthos biomass months 7-12, units kilometers 
7. Fall Northward Macrobenthos Center of Gravity = average northward location of macrobenthos biomass months 7-12, units kilometers 8. Fall Northward Macrobenthos Center of Gravity SE = standard error of average northward location of macrobenthos biomass months 7-12, units kilometers 
9. Spring Eastward Macrobenthos Center of Gravity = average eastward location of macrobenthos biomass months 1-6, units kilometers 10. Spring Eastward Macrobenthos Center of Gravity SE = standard error of average eastward location of macrobenthos biomass months 1-6, units kilometers 
11. Spring Northward Macrobenthos Center of Gravity = average northward location of macrobenthos biomass months 1-6, units kilometers 12. Spring Northward Macrobenthos Center of Gravity SE = standard error of average northward location of macrobenthos biomass months 1-6, units kilometers 
13. Spring Megabenthos Biomass Estimate = aggregate megabenthos biomass months 1-6, units relative grams per stomach 14. Spring Megabenthos Biomass Estimate SE = standard error of aggregate megabenthos biomass months 1-6, units relative grams per stomach 
15. Fall Megabenthos Biomass Estimate = aggregate megabenthos biomass months 7-12, units relative grams per stomach 16. Fall Megabenthos Biomass Estimate SE = standard error of aggregate megabenthos biomass months 7-12, units relative grams per stomach 
17. Fall Eastward Megabenthos Center of Gravity = average eastward location of megabenthos biomass months 7-12, units kilometers 18. Fall Eastward Megabenthos Center of Gravity SE = standard error of average eastward location of megabenthos biomass months 7-12, units kilometers 
19. Fall Northward Megabenthos Center of Gravity = average northward location of megabenthos biomass months 7-12, units kilometers 20. Fall Northward Megabenthos Center of Gravity SE = standard error of average northward location of megabenthos biomass months 7-12, units kilometers 
21. Spring Eastward Megabenthos Center of Gravity = average eastward location of megabenthos biomass months 1-6, units kilometers 22. Spring Eastward Megabenthos Center of Gravity SE = standard error of average eastward location of megabenthos biomass months 1-6, units kilometers 
23. Spring Northward Megabenthos Center of Gravity = average northward location of megabenthos biomass months 1-6, units kilometers 24. Spring Northward Megabenthos Center of Gravity SE = standard error of average northward location of megabenthos biomass months 1-6, units kilometers

```{r vars_benthos_index}
# Pull all var names
vars <- ecodata::benthos_index |>
   dplyr::select(Var, Units) |>
   dplyr::distinct()

DT::datatable(vars)
```
**Indicator Category**:

- [X] Extensive analysis, not yet published


## Public Availability

Source data are publicly available. All data and code available on GitHub at https://github.com/NOAA-EDAB/benthosindex

## Accessibility and Constraints

_No response_

**tech-doc link**
<https://noaa-edab.github.io/tech-doc/benthos_index.html>

