# Cetacean Weekly Acoustic Presence {#cetacean_acoustic}

**Description**: This figure shows a summary of weekly acoustic presence of eight cetacean species summarized across two years at four recording sites in southern New England.

**Indicator family**: 

- [X] Megafauna


**Contributor(s)**: Sofie Van Parijs, Annamaria DeAngelis, Tyler Aldrich, Rochelle Gordon, Amanda Holdman, Xavier Mouy, Tim Rowell, Sara Tennant, Annabel Westell, and Genevieve Davis

**Affiliations**: NEFSC

```{r echo=FALSE}
knitr::opts_chunk$set(echo = F)
library(ecodata)
```
## Introduction to Indicator
Passive acoustic monitoring (PAM) is a non-invasive method that provides continuous sampling effort in areas that are otherwise difficult to observe species presence and provides information on multiple species simultaneously. PAM allows for understanding trends in species distribution, habitat use, cumulative impacts and risk mitigation, and evaluating potential behavioral and distributional changes resulting from climate change or anthropogenic activity. 

This indicator examined the weekly acoustic presence of eight cetacean species summarized across two years. This summary allows for an improved understanding of how frequently each of the species/family were acoustically active and whether they exhibit seasonal patterns in their activity within the region. This data serves as a baseline summary prior to construction of the southern New England Offshore Wind Energy Areas and full description of the research is available in Van Parijs et al. 2023[@vanparijs_establishing_2023].

## Key Results and Visualizations
![Figure 1](https://github.com/user-attachments/assets/e3602f2a-a991-40c7-8511-110f878ec393){width=100%}


## Indicator statistics 
Spatial scale: A total of four PAM units were located on Cox Ledge and west of Nantucket shoals in southern New England near the boundaries of the MA-RI Offshore Wind Energy Areas. The Distance between recorders was a maximum of 100km East-West, and 50 km North-South.

Temporal scale: Acoustic data were summarized from a two-year continuous sampling period spanning 2020 - 2022.

**Synthesis Theme**:

- [X] Multiple System Drivers


```{r autostats_cetacean_acoustic}
# Either from Contributor or ecodata
```

## Implications
Acoustic summaries show that Southern New England is an important habitat for cetaceans, with eight species/families present for at least 9 months of the year. 

Winter Presence: Harbor porpoise, North Atlantic right whale, fin, and humpback whales all showed consistent presence throughout the winter months and a decrease in summer months. 

Summer Presence: Sperm whale acoustic presence was also seasonal in nature with a peak in late spring and summer (May to July) with a maximum of 4 d a week of presence. There was sporadic presence of between 1 and 2 d a week across all other months except for March when no acoustic activity was heard. 

Spring Presence: Sei whales showed a peak in late winter and early spring (February to May) with sporadic presence throughout all other months. 

Spring and Fall/Autumn Presence: Minke whales saw increased activity in the spring, with a maximum of 2 d a week of presence (April/May) and a prolonged increase in summer through fall (July to November) with a maximum of 4 d a week of call presences. 

Continuous Seasonal Presence: Delphinids showed no seasonal pattern and were acoustically active 7 d a week, with only a slight decrease in activity in March and April. 

Occasional Presence: Blue whales were only present on a few occasions in the winter (January/ February).

## Get the data

**Point of contact**: [Rebecca Van Hoeck (rebecca.vanhoeck@noaa.gov); Genevieve Davis (genevieve.davis@noaa.gov)](mailto:Rebecca Van Hoeck (rebecca.vanhoeck@noaa.gov); Genevieve Davis (genevieve.davis@noaa.gov)){.email}

**ecodata name**: `ecodata::cetacean_acoustic`

**Variable definitions**

Name: Acoustic presence; Definition: At least 1 species-specific call was observed on a given day. For North Atlantic right whale acoustic presence is defined as 3 species-specific calls observed on a given day.  
Name: Weekly acoustic presence; Definition: median number of days of acoustic presence per calendar week across all data at four recording sites

```{r vars_cetacean_acoustic}
# Pull all var names
vars <- ecodata::cetacean_acoustic |>
   dplyr::select(Var) |>
   dplyr::distinct()

DT::datatable(vars)
```
**Indicator Category**:

- [X] Published Methods


## Public Availability

Source data are publicly available.

## Accessibility and Constraints

_No response_

**tech-doc link**
<https://noaa-edab.github.io/tech-doc/cetacean_acoustic.html>

