# Chesapeake Bay Temperature {#ch_bay_temp}

**Description**: This data is collected from the CBIBS buoy system.

**Indicator family**: 

- [X] Oceanographic


**Contributor(s)**: Charles Pellerin

**Affiliations**: NEFSC

```{r echo=FALSE}
knitr::opts_chunk$set(echo = F)
library(ecodata)
```
## Introduction to Indicator
The indicator is the Water Temperature and the Water Salinity.

## Key Results and Visualizations
The salinity data, when compared to the 2010-2023 historical average reveals a distinct pattern. Salinity observations are lower than the historical averages from February to October and higher than historical averages during the fall (around October to November). This variation in salinity is likely due to increased rain events and freshwater input during spring and early summer. Spring 2024 salinity observations exceeded historical minimum salinity values, and remained below average throughout the summer. In early Fall, 2024 averages surpassed historical averages.

### MAB

```{r plot_ch_bay_tempMAB}
# Plot indicator
ggplotObject <- ecodata::plot_ch_bay_temp(report='MidAtlantic')
ggplotObject
```


## Indicator statistics 
Spatial scale: Main stem of the Chesapeake Bay

Temporal scale: Annual

**Synthesis Theme**:

- [X] Multiple System Drivers


```{r autostats_ch_bay_temp}
# Either from Contributor or ecodata
```

## Implications
Since 2007, the Chesapeake Bay Interpretive Buoy System (CBIBS) has provided invaluable, high-quality water quality data, underscoring the critical need for continuous monitoring of key parameters. The system has consistently collected robust data on meteorological, water quality and oceanographic parameters. While the system experienced maintenance challenges inherent in complex observational systems deployed in dynamic marine environments, the vast majority of the data collected is reliable and paints a picture of ongoing changes in Chesapeake Bay. This publicly available resource provides a high degree of confidence for Chesapeake Bay stakeholders, and is a vital tool for informing evidence-based environmental management decisions. The long-term nature of the CBIBS dataset allows for the detection of long-term trends in water column habitat and the assessment of restoration efforts.  

Salinity averages during the 2024 season were consistently lower than historical averages from 2008 - 2023 across all locations. Lower salinity during much of the summer was unexpected because much of the region was experiencing drier-than-normal weather. Despite this, the salinity levels generally remained within the historical range, with the exception of Annapolis and Gooses Reef. This shift in salinity could have important implications for living resources in the bay, particularly for species that are sensitive to changes in salinity, such as oysters, crabs, and various fish species. Lower salinity in the spring was likely driven by higher precipitation and flow. Years with high flow are typically better for striped bass spawning, however, the annual surveys of juvenile striped bass in [Maryland](https://news.maryland.gov/dnr/2024/10/17/results-of-chesapeake-bay-2024-young-of-year-striped-bass-survey-show-little-change/) and [Virginia](https://www.vims.edu/newsandevents/topstories/2024/juvenile-striped-bass-survey.php) showed juvenile indexes below the long-term average. Generally, low salinities in spring and summer can suppress oyster growth and reproduction (number of surviving spat) both in the wild and at hatcheries. Oyster survey data was not available at the time this report was drafted so effects on oysters cannot be correlated to the salinity observations at this time. Low salinity across the Chesapeake Bay allows invasive blue catfish to inhabit larger areas in the Bay. Invasive blue catfish prey on striped bass, blue crabs and other important finfish. 

Surface water temperatures during 2024 were mostly at or above historical averages across all sites, with notably above average temperatures during spring and summer months, and greater instances of at or below historical average temperatures towards the end of the year. The above average temperatures observed in the spring and summer months, may have had important effects on key living resources, such as striped bass, oysters, and other fish species within Chesapeake Bay. Above average temperatures could lead to decreases in the area of suitable habitat, disruptions in migration patterns, increased stress and mortality of aquatic species, and other negative impacts on key living resources within the Chesapeake Bay. Specifically, water temperatures above 82.4°F exceed suitable habitat thresholds for striped bass from about mid-July to mid-August at all buoy locations where data was available. Water temperatures coupled with [low dissolved oxygen](https://www.vims.edu/newsandevents/topstories/2024/2024-dead-zone-report-card.php) during spring and summer exacerbate impacts on habitat suitability for striped bass, blue crabs and other species. A number of anglers reported good catches of red drum during the summer. During this summer’s routine sampling at Poplar Island, NOAA Chesapeake Bay Office scientists recorded the highest number of red drum since monitoring began in 1995. This may indicate habitat conditions in the lower and upper Bay are conducive to red drum. This may be related to rising water temperatures linked to climate change as well as conservation measures implemented by management agencies. However, there are few new studies focused on understanding these changes.

At the end of 2024, CBIBS completed its second successful year of monitoring hypoxia in Chesapeake Bay, increasing from three locations in 2023 to six locations in 2024. This additional data source is the first of its kind in the region as it provides detailed insights into dissolved oxygen and other key water parameters throughout the entire water column. Data is collected every 10 minutes and will allow users to understand how the water column fluctuates throughout the year. This data will be invaluable to understand long term water quality trends at fixed depths, and how various fluctuations in key water quality parameters, especially dissolved oxygen, impact living resources within Chesapeake Bay. While the project's results are not provided for the 2024 state of the ecosystem report, they will be released in early spring 2025. This release will be accompanied by a standalone report detailing the data collection methodology and the quality assurance protocols implemented to ensure data quality.

## Get the data

**Point of contact**: [Charles Pellerin (charles.pellerin@noaa.gov)](mailto:Charles Pellerin (charles.pellerin@noaa.gov)){.email}

**ecodata name**: `ecodata::ch_bay_temp`

**Variable definitions**

1) Temperature; degree C

```{r vars_ch_bay_temp}
# Pull all var names
vars <- ecodata::ch_bay_temp |>
   dplyr::select(Var, Units) |>
   dplyr::distinct()

DT::datatable(vars)
```
**Indicator Category**:

- [X] Database pull


## Public Availability

Source data are publicly available.

## Accessibility and Constraints

_No response_

**tech-doc link**
<https://noaa-edab.github.io/tech-doc/ch_bay_temp.html>

