# Chesapeake Bay Seasonal Sea Surface Temperature Anomaly {#ches_bay_sst}

**Description**: Chesapeake Bay Seasonal Sea Surface Temperature Anomaly

**Indicator family**: 

- [X] Oceanographic


**Contributor(s)**: Ron Vogel, Bruce Vogt

**Affiliations**: University of Maryland, Cooperative Institute for Satellite Earth System Studies

```{r echo=FALSE}
knitr::opts_chunk$set(echo = F)
library(ecodata)
```
## Introduction to Indicator
Seasonal spatial anomaly maps represent the difference between a seasonal average sea surface temperature (SST) and a long-term average SST for that season. For the Chesapeake Bay, the long-term average is derived from the SST time series from 2007 to the year immediately prior to the current year (max(Year) - 1). This reference period serves as a benchmark for comparing current observations. Hence, the anomaly represents the degree to which the current seasonal average departs from historical average, either colder or warmer, indicating whether the current temperature conditions may be favorable or unfavorable for marine species.

## Key Results and Visualizations
For 2024, winter SST’s show conditions to be roughly 1 – 1.5 degrees Celsius warmer than the prior 17-year average winter (2007-2023) throughout Chesapeake Bay. Spring conditions are similar: roughly 0.5 – 1.5 degrees Celsius warmer than average. Summer data are not consistent with Chesapeake Bay buoy data (when comparing to long-term averages), so those values are not shown and are currently under investigation. Fall conditions are mixed spatially but do not exceed 0.5 degrees Celsius colder or warmer.

```{r plot_ches_bay_sstMAB}
# Plot indicator
ggplotObject <- ecodata::plot_ches_bay_sst(report='MidAtlantic')
ggplotObject
```


## Indicator statistics 
Spatial scale: Data from two satellite instruments, AVHRR at 1 km spatial resolution and VIIRS at 750 m spatial resolution, are co-gridded to an 830 m spatial grid. Overpasses from the two instruments on all current operational satellites are composited into a daily scene in order to maximize geographic coverage on a per-day basis, i.e. minimize data gaps from clouds. Seasonal averaging further increases geographic coverage.

Temporal scale: Only nighttime satellite overpasses are used in the seasonal averages, i.e. the data do not represent daytime solar heating of the water surface. Seasons for Chesapeake Bay are Dec-Feb (winter), Mar-May (spring), Jun-Aug (summer), and Sep-Nov (fall).

**Synthesis Theme**:

- [X] Multiple System Drivers
- [X] Regime Shifts
- [X] Ecosystem Reorganization


```{r autostats_ches_bay_sst}
# Either from Contributor or ecodata
```

## Implications
The warm water during the 2024 winter (Dec 2023 – Feb 2024) was likely favorable to blue crabs (Callinectes sapidus) by reducing their overwintering mortality. Followed by a warm spring, the blue crabs may have benefited from an extended growing season (warm water earlier in the spring), but striped bass may be experiencing a shortening of their spawning season. In summer, water temperature exceeded typical values for the early part of the summer (based on buoy data, satellite data not shown here), and fall temperatures, with roughly average conditions, may have benefited most of the tracked fishery species.

## Get the data

**Point of contact**: [Ron Vogel (ronald.vogel@noaa.gov)](mailto:Ron Vogel (ronald.vogel@noaa.gov)){.email}

**ecodata name**: `ecodata::ches_bay_sst`

**Variable definitions**

1) sst: sea surface temperature 2023, Celsius 
2) sst_climatol: sea surface temperature climatology 2007-2022, Celsius 
3) sst_anomaly: sea surface temperature anomaly 2023 minus 2007-2022, Celsius

```{r vars_ches_bay_sst}
# Pull all var names
vars <- ecodata::ches_bay_sst |>
   dplyr::select(Var) |>
   dplyr::distinct()

DT::datatable(vars)
```
**Indicator Category**:

- [X] Published Methods
- [X] Database pull with analysis


## Public Availability

Source data are NOT publicly available.

## Accessibility and Constraints

The seasonal SST anomaly data files (including the SST long-term climatology) are available from Ron Vogel at ronald.vogel@noaa.gov. The time series of daily SST, seasonal average SST, and other time intervals for 2007 – present, are available to the public at: https://www.star.nesdis.noaa.gov/pub/socd1/ecn/data/avhrr-viirs/sst-ngt. For more information about this SST data set, see: https://eastcoast.coastwatch.noaa.gov/cw_avhrr-viirs_sst.php. For other inquiries about this data set, contact Ron Vogel at ronald.vogel@noaa.gov.

**tech-doc link**
<https://noaa-edab.github.io/tech-doc/ches_bay_sst.html>

