# Phytoplankton {#chl_pp}

**Description**: Satellite derived phytoplankton data including chlorophyll concentration, primary production, and phytoplankton size class for the Northeast Continental Shelf and ecological production units.

**Indicator family**: 

- [X] Lower trophic levels


**Contributor(s)**: Kimberly Hyde

**Affiliations**: NEFSC

```{r echo=FALSE}
knitr::opts_chunk$set(echo = F)
library(ecodata)
```
## Introduction to Indicator


## Key Results and Visualizations
The seasonal cycle of phytoplankton biomass typically includes peaks during the winter-spring and fall seasons and minimums during the summer months, however exact bloom timing varies by region. Primary productivity, increases during bloom periods due to increased biomass, but peaks during the summer months when daylight and temperatures are greatest. In 2024, concentrations of chlorophyll _a_ (a proxy for biomass) were above average during the winter-spring and fall bloom periods in all regions (figure 2.2.1). There was no evidence of a repeat of the widespread _Tripos muelleri_ bloom [add link to the 2023 highlights catalog] observed in 2023 (ADD REFERENCE TO THE 2023 NE Report). There was, however, an atypical localized coccolithophore bloom [once available, add link to the coccolithophore section of the 2024 highlights catalog] south of Long Island, New York in late June to early July. 

[INSERT MAB, GB and GOM 2024 chlorophyll plots as Figure 2.2.1]

[Caption: 2024 weekly median chlorophyll a concentrations (green line) for the Mid-Atlantic, Georges Bank and Gulf of Maine ecoregions. The black line is the climatological mean (1998-2020) and the shaded region represents ???] 

Mid-Atlantic primary production peaked in early summer in 2024 and was above average in late fall, which is attributed to the above average late fall bloom (figures 2.2.2a). Georges Bank fall productivity was also above average due to the abnormally large fall phytoplankton bloom (figure 2.2.2b). Conversely, the Gulf of Maine spring bloom accounted for the largest above average spike in that region (figure 2.2.2c).

[INSERT MAB, GB and GOM 2024 PP plots as Figure 2.2.2]

[Caption: 2024 weekly median primary productivity (green line) for the Mid-Atlantic, Georges Bank and Gulf of Maine ecoregions. The black line is the climatological mean (1998-2020) and the shaded region represents ???] 

The seasonal cycles of phytoplankton size distribution are typically dominated by larger-celled microplankton (> 20 um) during the winter-spring and fall bloom periods, while smaller-celled nanoplankton (0.2-20 um) dominate during the warmer summer months. The smallest phytoplankton group, picoplankton (<0.2 um), represent the smallest fraction of the phytoplankton community in the region. Phytoplankton size class distributions were near average for most of the year, except during the early fall bloom. The 2024 phytoplankton size class distribution was near the seasonal climatology in all regions except during the above average phytoplankton bloom periods (figure 2.2.3).

[INSERT MAB, GB and GOM 2024 PSC plots as figure 2.2.3]

[Caption: The climatological (1998-2020) and 2024 seasonal cycle of phytoplankton size class percent composition. The green, orange, and purple shaded regions represents the weekly climatological percent composition of microplankton (phytoplankton greater than 20 um), nanoplankton (phytoplankton between 0.2 and 20 um) and picoplankton (phytoplankton less than 0.2 um) respectively. The solid green and orange lines are the 2024 weekly mean percent microplankton and nanoplankton respectively. 




There is high interannual variability of the seasonal phytoplankton cycle.  At the monthly scale, MAB chlorophyll and primary production are increasing during January and there has been a decrease in September chlorophyll, likely due to extension of the [summer stratification](https://noaa-edab.github.io/catalog/transition-dates.html) and delayed fall turnover. Fall and winter chlorophyll and primary production are increasing on Georges Bank and Gulf of Maine.

### MidAtlantic


```{r plot_chl_ppMidAtlanticchlweeklyMAB}
# Plot indicator
ggplotObject <- ecodata::plot_chl_pp(report= 'MidAtlantic', EPU = 'MAB', varName= 'chl', plottype = 'weekly',n=10)
ggplotObject
```


```{r plot_chl_ppMidAtlanticchlmonthlyMAB}
# Plot indicator
ggplotObject <- ecodata::plot_chl_pp(report= 'MidAtlantic', EPU = 'MAB', varName= 'chl', plottype = 'monthly',n=10)
ggplotObject
```


```{r plot_chl_ppMidAtlanticppweeklyMAB}
# Plot indicator
ggplotObject <- ecodata::plot_chl_pp(report= 'MidAtlantic', EPU = 'MAB', varName= 'pp', plottype = 'weekly',n=10)
ggplotObject
```


```{r plot_chl_ppMidAtlanticppmonthlyMAB}
# Plot indicator
ggplotObject <- ecodata::plot_chl_pp(report= 'MidAtlantic', EPU = 'MAB', varName= 'pp', plottype = 'monthly',n=10)
ggplotObject
```


```{r plot_chl_ppMidAtlanticsizeweeklyMAB}
# Plot indicator
ggplotObject <- ecodata::plot_chl_pp(report= 'MidAtlantic', EPU = 'MAB', varName= 'size', plottype = 'weekly',n=10)
ggplotObject
```


```{r plot_chl_ppMidAtlanticsizemonthlyMAB}
# Plot indicator
ggplotObject <- ecodata::plot_chl_pp(report= 'MidAtlantic', EPU = 'MAB', varName= 'size', plottype = 'monthly',n=10)
ggplotObject
```

### NewEngland


```{r plot_chl_ppNewEnglandchlweeklyGB}
# Plot indicator
ggplotObject <- ecodata::plot_chl_pp(report= 'NewEngland', EPU = 'GB', varName= 'chl', plottype = 'weekly',n=10)
ggplotObject
```

```{r plot_chl_ppNewEnglandchlweeklyGOM}
# Plot indicator
ggplotObject <- ecodata::plot_chl_pp(report= 'NewEngland', EPU = 'GOM', varName= 'chl', plottype = 'weekly',n=10)
ggplotObject
```


```{r plot_chl_ppNewEnglandchlmonthlyGB}
# Plot indicator
ggplotObject <- ecodata::plot_chl_pp(report= 'NewEngland', EPU = 'GB', varName= 'chl', plottype = 'monthly',n=10)
ggplotObject
```

```{r plot_chl_ppNewEnglandchlmonthlyGOM}
# Plot indicator
ggplotObject <- ecodata::plot_chl_pp(report= 'NewEngland', EPU = 'GOM', varName= 'chl', plottype = 'monthly',n=10)
ggplotObject
```


```{r plot_chl_ppNewEnglandppweeklyGB}
# Plot indicator
ggplotObject <- ecodata::plot_chl_pp(report= 'NewEngland', EPU = 'GB', varName= 'pp', plottype = 'weekly',n=10)
ggplotObject
```

```{r plot_chl_ppNewEnglandppweeklyGOM}
# Plot indicator
ggplotObject <- ecodata::plot_chl_pp(report= 'NewEngland', EPU = 'GOM', varName= 'pp', plottype = 'weekly',n=10)
ggplotObject
```


```{r plot_chl_ppNewEnglandppmonthlyGB}
# Plot indicator
ggplotObject <- ecodata::plot_chl_pp(report= 'NewEngland', EPU = 'GB', varName= 'pp', plottype = 'monthly',n=10)
ggplotObject
```

```{r plot_chl_ppNewEnglandppmonthlyGOM}
# Plot indicator
ggplotObject <- ecodata::plot_chl_pp(report= 'NewEngland', EPU = 'GOM', varName= 'pp', plottype = 'monthly',n=10)
ggplotObject
```


```{r plot_chl_ppNewEnglandsizeweeklyGB}
# Plot indicator
ggplotObject <- ecodata::plot_chl_pp(report= 'NewEngland', EPU = 'GB', varName= 'size', plottype = 'weekly',n=10)
ggplotObject
```

```{r plot_chl_ppNewEnglandsizeweeklyGOM}
# Plot indicator
ggplotObject <- ecodata::plot_chl_pp(report= 'NewEngland', EPU = 'GOM', varName= 'size', plottype = 'weekly',n=10)
ggplotObject
```


```{r plot_chl_ppNewEnglandsizemonthlyGB}
# Plot indicator
ggplotObject <- ecodata::plot_chl_pp(report= 'NewEngland', EPU = 'GB', varName= 'size', plottype = 'monthly',n=10)
ggplotObject
```

```{r plot_chl_ppNewEnglandsizemonthlyGOM}
# Plot indicator
ggplotObject <- ecodata::plot_chl_pp(report= 'NewEngland', EPU = 'GOM', varName= 'size', plottype = 'monthly',n=10)
ggplotObject
```


## Indicator statistics 
Spatial scale: By EPU and gridded for the entire shelf

Temporal scale: Daily, weekly, monthly, annual, climatology (1998 to current year)

**Synthesis Theme**:

- [X] Multiple System Drivers


```{r autostats_chl_pp}
# Either from Contributor or ecodata
```

## Implications
Phytoplankton abundance, productivity, diversity, cell size, phenology, and carbon fluxes are regulated by the local physical and chemical environment and grazing. Interannual and climatological changes in temperature, freshwater inputs (due to ice sheet melting and/or enhanced river discharge), wind direction, and wind speed can alter the circulation patterns, upwelling conditions, and nutrient fluxes, directly affecting the timing, location, species composition of phytoplankton blooms in the NES. As the NES responds to warming, changing phenologies, changing chemistry, and changes in circulation patterns, we must understand how varying biophysical interactions control phytoplankton and subsequently affect fisheries, their habitats and the people, businesses and communities that depend on them.

## Get the data

**Point of contact**: [kimberly.hyde@noaa.gov](mailto:kimberly.hyde@noaa.gov){.email}

**ecodata name**: `ecodata::chl_pp`

**Variable definitions**

1) Chlorophyll; Concentration of chlorophyll _a_ in the near surface (first euphotic depth) waters; mg m^-3 
2) Chlorophyll anomaly; Ratio of chlorophyll _a_ concentration to the long-term (1998 to present year) climatology; unitless (ratio) 
3) Primary productivity; Daily net primary production of biomass expressed as carbon per unit volume in seawater per day; gCarbon m^-2 d^-1 
4) Primary productivity anomaly;  Ratio of net primary production to the long-term (1998 to present year) climatology; unitless (ratio) 
5) Microplankton chlorophyll; Concentration of chlorophyll _a_ in the near surface (first euphotic depth) waters from the microplankton (20-200 µm) fraction; mg m^-3 
6) Nanoplankton chlorophyll; Concentration of chlorophyll _a_ in the near surface (first euphotic depth) waters from the nanoplankton (2-20 µm) fraction; mg m^-3 
7) Picoplankton chlorophyll; Concentration of chlorophyll _a_ in the near surface (first euphotic depth) waters from the picoplankton (0.2-2 µm) fraction; mg m^-3 
8) Annual summed primary production - TBD

```{r vars_chl_pp}
# Pull all var names
vars <- ecodata::chl_pp |>
   dplyr::select(Var, Units) |>
   dplyr::distinct()

DT::datatable(vars)
```
**Indicator Category**:

- [X] Published Methods
- [X] Other


**Indicator Category**:

Publicly available satellite data that are processed and analyzed

## Public Availability

Source data are publicly available.

## Accessibility and Constraints

_No response_

**tech-doc link**
<https://noaa-edab.github.io/tech-doc/chl_pp.html>

