# Comdat Profitability Indices {#comdat_profit}

```{r echo=FALSE}
knitr::opts_chunk$set(echo = F)
library(ecodata)
library(downloadthis)
```

**Description**: Coming soon

**Indicator family**: 

- [x] Economic


**Contributor(s)**: Coming soon

**Affiliations**: NEFSC

```{r download_comdat_profit}
# Download dataset from ecodata
downloadthis::download_this(
ecodata::comdat_profit,
output_name = 'comdat_profit',
output_extension = '.csv',
button_label = 'Download dataset from `ecodata`',
button_type = 'default',
has_icon = TRUE,
icon = 'fa fa-save',
class = 'hvr-sweep-to-left',
csv2 = F)
```

## Introduction to Indicator
Coming soon

## Key Results and Visualizations
Coming soon

**all, Mid-Atlantic Bight**
```{r plot_comdat_profitallMidAtlanticMAB}
# Plot indicator
ggplotObject <- ecodata::plot_comdat_profit(varName = 'all', report = 'MidAtlantic', EPU = 'MAB', n = 10)
ggplotObject
try(downloadthis::download_this(
ggplotObject$data,
output_name = 'comdat_profitallMidAtlanticMAB',
output_extension = '.csv',
button_label = 'Download the data plotted above',
button_type = 'default',
has_icon = TRUE,
icon = 'fa fa-save',
class = 'hvr-sweep-to-left',
csv2 = F),
silent = TRUE)
```

**profit_index, Mid-Atlantic Bight**
```{r plot_comdat_profitprofit_indexMidAtlanticMAB}
# Plot indicator
ggplotObject <- ecodata::plot_comdat_profit(varName = 'profit_index', report = 'MidAtlantic', EPU = 'MAB', n = 10)
ggplotObject
try(downloadthis::download_this(
ggplotObject$data,
output_name = 'comdat_profitprofit_indexMidAtlanticMAB',
output_extension = '.csv',
button_label = 'Download the data plotted above',
button_type = 'default',
has_icon = TRUE,
icon = 'fa fa-save',
class = 'hvr-sweep-to-left',
csv2 = F),
silent = TRUE)
```

**cost_index, Mid-Atlantic Bight**
```{r plot_comdat_profitcost_indexMidAtlanticMAB}
# Plot indicator
ggplotObject <- ecodata::plot_comdat_profit(varName = 'cost_index', report = 'MidAtlantic', EPU = 'MAB', n = 10)
ggplotObject
try(downloadthis::download_this(
ggplotObject$data,
output_name = 'comdat_profitcost_indexMidAtlanticMAB',
output_extension = '.csv',
button_label = 'Download the data plotted above',
button_type = 'default',
has_icon = TRUE,
icon = 'fa fa-save',
class = 'hvr-sweep-to-left',
csv2 = F),
silent = TRUE)
```

**revenue_index, Mid-Atlantic Bight**
```{r plot_comdat_profitrevenue_indexMidAtlanticMAB}
# Plot indicator
ggplotObject <- ecodata::plot_comdat_profit(varName = 'revenue_index', report = 'MidAtlantic', EPU = 'MAB', n = 10)
ggplotObject
try(downloadthis::download_this(
ggplotObject$data,
output_name = 'comdat_profitrevenue_indexMidAtlanticMAB',
output_extension = '.csv',
button_label = 'Download the data plotted above',
button_type = 'default',
has_icon = TRUE,
icon = 'fa fa-save',
class = 'hvr-sweep-to-left',
csv2 = F),
silent = TRUE)
```

**all, Georges Bank**
```{r plot_comdat_profitallNewEnglandGB}
# Plot indicator
ggplotObject <- ecodata::plot_comdat_profit(varName = 'all', report = 'NewEngland', EPU = 'GB', n = 10)
ggplotObject
try(downloadthis::download_this(
ggplotObject$data,
output_name = 'comdat_profitallNewEnglandGB',
output_extension = '.csv',
button_label = 'Download the data plotted above',
button_type = 'default',
has_icon = TRUE,
icon = 'fa fa-save',
class = 'hvr-sweep-to-left',
csv2 = F),
silent = TRUE)
```

**profit_index, Georges Bank**
```{r plot_comdat_profitprofit_indexNewEnglandGB}
# Plot indicator
ggplotObject <- ecodata::plot_comdat_profit(varName = 'profit_index', report = 'NewEngland', EPU = 'GB', n = 10)
ggplotObject
try(downloadthis::download_this(
ggplotObject$data,
output_name = 'comdat_profitprofit_indexNewEnglandGB',
output_extension = '.csv',
button_label = 'Download the data plotted above',
button_type = 'default',
has_icon = TRUE,
icon = 'fa fa-save',
class = 'hvr-sweep-to-left',
csv2 = F),
silent = TRUE)
```

**cost_index, Georges Bank**
```{r plot_comdat_profitcost_indexNewEnglandGB}
# Plot indicator
ggplotObject <- ecodata::plot_comdat_profit(varName = 'cost_index', report = 'NewEngland', EPU = 'GB', n = 10)
ggplotObject
try(downloadthis::download_this(
ggplotObject$data,
output_name = 'comdat_profitcost_indexNewEnglandGB',
output_extension = '.csv',
button_label = 'Download the data plotted above',
button_type = 'default',
has_icon = TRUE,
icon = 'fa fa-save',
class = 'hvr-sweep-to-left',
csv2 = F),
silent = TRUE)
```

**revenue_index, Georges Bank**
```{r plot_comdat_profitrevenue_indexNewEnglandGB}
# Plot indicator
ggplotObject <- ecodata::plot_comdat_profit(varName = 'revenue_index', report = 'NewEngland', EPU = 'GB', n = 10)
ggplotObject
try(downloadthis::download_this(
ggplotObject$data,
output_name = 'comdat_profitrevenue_indexNewEnglandGB',
output_extension = '.csv',
button_label = 'Download the data plotted above',
button_type = 'default',
has_icon = TRUE,
icon = 'fa fa-save',
class = 'hvr-sweep-to-left',
csv2 = F),
silent = TRUE)
```

**all, Gulf of Maine**
```{r plot_comdat_profitallNewEnglandGOM}
# Plot indicator
ggplotObject <- ecodata::plot_comdat_profit(varName = 'all', report = 'NewEngland', EPU = 'GOM', n = 10)
ggplotObject
try(downloadthis::download_this(
ggplotObject$data,
output_name = 'comdat_profitallNewEnglandGOM',
output_extension = '.csv',
button_label = 'Download the data plotted above',
button_type = 'default',
has_icon = TRUE,
icon = 'fa fa-save',
class = 'hvr-sweep-to-left',
csv2 = F),
silent = TRUE)
```

**profit_index, Gulf of Maine**
```{r plot_comdat_profitprofit_indexNewEnglandGOM}
# Plot indicator
ggplotObject <- ecodata::plot_comdat_profit(varName = 'profit_index', report = 'NewEngland', EPU = 'GOM', n = 10)
ggplotObject
try(downloadthis::download_this(
ggplotObject$data,
output_name = 'comdat_profitprofit_indexNewEnglandGOM',
output_extension = '.csv',
button_label = 'Download the data plotted above',
button_type = 'default',
has_icon = TRUE,
icon = 'fa fa-save',
class = 'hvr-sweep-to-left',
csv2 = F),
silent = TRUE)
```

**cost_index, Gulf of Maine**
```{r plot_comdat_profitcost_indexNewEnglandGOM}
# Plot indicator
ggplotObject <- ecodata::plot_comdat_profit(varName = 'cost_index', report = 'NewEngland', EPU = 'GOM', n = 10)
ggplotObject
try(downloadthis::download_this(
ggplotObject$data,
output_name = 'comdat_profitcost_indexNewEnglandGOM',
output_extension = '.csv',
button_label = 'Download the data plotted above',
button_type = 'default',
has_icon = TRUE,
icon = 'fa fa-save',
class = 'hvr-sweep-to-left',
csv2 = F),
silent = TRUE)
```

**revenue_index, Gulf of Maine**
```{r plot_comdat_profitrevenue_indexNewEnglandGOM}
# Plot indicator
ggplotObject <- ecodata::plot_comdat_profit(varName = 'revenue_index', report = 'NewEngland', EPU = 'GOM', n = 10)
ggplotObject
try(downloadthis::download_this(
ggplotObject$data,
output_name = 'comdat_profitrevenue_indexNewEnglandGOM',
output_extension = '.csv',
button_label = 'Download the data plotted above',
button_type = 'default',
has_icon = TRUE,
icon = 'fa fa-save',
class = 'hvr-sweep-to-left',
csv2 = F),
silent = TRUE)
```


## Indicator statistics 
Spatial scale: Coming soon

Temporal scale: Coming soon

**Synthesis Theme**:




```{r autostats_comdat_profit}
# Either from Contributor or ecodata
```

## Implications
Coming soon

## Get the data

**Point of contact**: [EDAB (edab.data@noaa.gov)](mailto:EDAB (edab.data@noaa.gov)){.email}

**ecodata name**: `ecodata::comdat_profit`

**Variable definitions**

Coming soon

```{r vars_comdat_profit}
# Pull all var names
vars <- ecodata::comdat_profit |>
   dplyr::select(Var) |>
   dplyr::distinct()

DT::datatable(vars)
```
**Indicator Category**:




## Public Availability

Source data are publicly available.

## Accessibility and Constraints

_No response_

**tech-doc link**
<https://noaa-edab.github.io/tech-doc/comdat_profit.html>

