# Dissolved Oxygen {#dissolved_oxygen}

```{r echo=FALSE}
knitr::opts_chunk$set(echo = F)
library(ecodata)
library(downloadthis)
```

*Last updated April 06, 2026*

**Description**: Coming soon

**Contributor(s)**: Coming soon

**Affiliations**: Coming soon

**Indicator Family**: 

- [x] Oceanographic
- [x] Habitat


**Indicator Category**:




**Synthesis Theme**:

- [x] Multiple System Drivers


```{r download_dissolved_oxygen}
# Download dataset from ecodata
downloadthis::download_this(
ecodata::dissolved_oxygen,
output_name = 'dissolved_oxygen',
output_extension = '.csv',
button_label = 'Download dataset from `ecodata`',
button_type = 'default',
has_icon = TRUE,
icon = 'fa fa-save',
class = 'hvr-sweep-to-left',
csv2 = F)
```

## Introduction to Indicator
Coming soon

## Key Results and Visualizations
Coming soon

****
```{r plot_dissolved_oxygen}
# Plot indicator
ggplotObject <- ecodata::plot_dissolved_oxygen()
ggplotObject
try(downloadthis::download_this(
ggplotObject$data,
output_name = 'dissolved_oxygen',
output_extension = '.csv',
button_label = 'Download the data plotted above',
button_type = 'default',
has_icon = TRUE,
icon = 'fa fa-save',
class = 'hvr-sweep-to-left',
csv2 = F),
silent = TRUE)
```
<br>

## Implications
Coming soon

## Indicator statistics 
Spatial scale: Coming soon

Temporal scale: Coming soon

**Variable definitions**

```{r vars_dissolved_oxygen}
# Pull all var names
vars <- ecodata::dissolved_oxygen |>
   dplyr::select(Var) |>
   dplyr::distinct()

DT::datatable(vars)
```
## Get the data

**Point of contact**: [Coming soon](mailto:Coming soon){.email}

**ecodata dataset**: `ecodata::dissolved_oxygen`

**Tech Doc link**: <https://noaa-edab.github.io/tech-doc/dissolved_oxygen.html>

### Public Availability

Source data are publicly available.

