# Highly Migratory Species Landings {#hms_landings}

**Description**: Aggregated Atlantic HMS landings data prepared for the Fisheries of the United States (FUS) report, spanning 2015-2023.

**Indicator family**: 

- [X] Megafauna


**Contributor(s)**: Heather Baertlein, Jackie Wilson, George Silva, Jennifer Cudney

**Affiliations**: NEFSC

```{r echo=FALSE}
knitr::opts_chunk$set(echo = F)
library(ecodata)
```
## Introduction to Indicator
This indicator provides a comprehensive summary of commercial landings information for HMS fisheries that target or retain tunas, swordfish, and sharks. These data can be evaluated independently, and can also be included in overall estimates of commercial fisheries landings in the northeast region.

## Key Results and Visualizations
Data are provided to SOE editors to generate the visualizations.

### MidAtlantic

```{r plot_hms_landingsMidAtlanticLandings}
# Plot indicator
ggplotObject <- ecodata::plot_hms_landings(report= 'MidAtlantic', varName= 'Landings')
ggplotObject
```

```{r plot_hms_landingsMidAtlanticRevenue}
# Plot indicator
ggplotObject <- ecodata::plot_hms_landings(report= 'MidAtlantic', varName= 'Revenue')
ggplotObject
```

### NewEngland

```{r plot_hms_landingsNewEnglandLandings}
# Plot indicator
ggplotObject <- ecodata::plot_hms_landings(report= 'NewEngland', varName= 'Landings')
ggplotObject
```

```{r plot_hms_landingsNewEnglandRevenue}
# Plot indicator
ggplotObject <- ecodata::plot_hms_landings(report= 'NewEngland', varName= 'Revenue')
ggplotObject
```


## Indicator statistics 
Spatial scale: Regionally by EPU

Temporal scale: Annual

**Synthesis Theme**:

- [X] Regime Shifts


```{r autostats_hms_landings}
# Either from Contributor or ecodata
```

## Implications
HMS landings data are summarized and presented in the annual HMS Stock Assessment and Fishery Evaluation (SAFE) report.

In 2021 the International Commission for the Conservation of Atlantic Tunas (ICCAT) finalized recommendations for a two-year retention ban for shortfin mako (ICCAT Rec. 21-09), which will also affect total overall landings of pelagic sharks in coming years.

## Get the data

**Point of contact**: [Jennifer Cudney (jennifer.cudney@noaa.gov)](mailto:Jennifer Cudney (jennifer.cudney@noaa.gov)){.email}

**ecodata name**: `ecodata::hms_landings`

**Variable definitions**

Year: data are summarized/aggregated by year. EPU: region, Mid Atlantic Bight (MAB) or New England (NE). HMS_Group: Management group for HMS, as defined below in “Data Processing”. Var: description of data, whether total landings or revenue. Units: description of unit of measure for data. Value: Represents either metric tons of landings or dollar value of landings by year and region.

```{r vars_hms_landings}
# Pull all var names
vars <- ecodata::hms_landings |>
   dplyr::select(Var) |>
   dplyr::distinct()

DT::datatable(vars)
```
**Indicator Category**:

- [X] Database pull with analysis


## Public Availability

Source data are NOT publicly available.

## Accessibility and Constraints

Confidential data may be released in “any aggregate or summary form which does not directly or indirectly disclose the identity or business of any person who submits such information” (Magnuson-Stevens Act § 402(b)(3); 16 U.S.C. 1881a(b)(3)). NOAA Fisheries presents such information only if it can be aggregated or summarized at a temporal and/or spatial level to maintain confidentiality of individuals, businesses, and related business information. Data that has been properly screened to meet data standards and confidentiality are publicly available via the Fisheries of the United States (FUS) landings portal. Canadian landings information, which are included in this analysis, are not included in the FUS portal (https://www.fisheries.noaa.gov/national/sustainable-fisheries/commercial-fisheries-landings). Data should be aggregated to protect data confidentiality (raw data are lumped by year and species). Data were included in the final summary unless there were 2 or less dealers involved in the landings, 2 or less vessels, or if any one dealer or vessel purchased/landed more than 2/3 of the total amount for that data unit. Please email jennifer.cudney@noaa.gov for further information and queries of source data.

**tech-doc link**
<https://noaa-edab.github.io/tech-doc/hms_landings.html>

