# Zooplankton Center of Gravity {#New Indicator}

**Description**: Model-based annual center of gravity indices for zooplankton groups sampled by NEFSC surveys

**Indicator family**: 

- [X] Lower trophic levels


**Contributor(s)**: Sarah Gaichas and Harvey Walsh

**Affiliations**: NEFSC

```{r echo=FALSE}
knitr::opts_chunk$set(echo = F)
library(ecodata)
```
## Introduction to Indicator
Zooplankton are primary consumers in marine ecosystems that transfer energy from phytoplankton to fish, marine mammals, and birds. The Northeast Fisheries Science Center has conducted [zooplankton surveys](https://www.fisheries.noaa.gov/inport/item/9286) since the 1970s. Spatially explicit indices of abundance for several zooplankton groups were estimated using spatio-temporal modeling.

The groups include:

**Copepods**

* _Calanus finmarchicus_, = Large copeopds SOE (used in small-large index)

* Large copepods ALL: _Calanus finmarchicus_, _Metridia lucens_, _Calanus minor_, _Eucalanus_ spp., _Calanus_ spp.

* Small copepods ALL: _Centropages typicus_, _Pseudocalanus_ spp., T_emora longicornis_, _Centropages hamatus_, _Paracalanus parvus_, _Acartia_ spp., _Clausocalanus arcuicornis_, _Acartia longiremis_, _Clausocalanus furcatus_, _Temora stylifera_, _Temora_ spp., _Tortanus discaudatus_, _Paracalanus_ spp.

* Small copeopods SOE (used in small-large index): _Centropages typicus_, _Pseudocalanus_ spp., _Temora longicornis_, _Centropages hamatus_

**Euphausiids** 

**Zooplankton volume**

## Key Results and Visualizations
Small copepods in fall are trending northeast similar to forage fish and aggregated survey species. Small copepods in spring are trending north. Large copepods in fall are trending west, similar to benthos. No significant change for Calfin, Euphausiids, or Zooplankton volume center of gravity over time.


## Indicator statistics 
Spatial scale: Full shelf

Temporal scale: Spring (January-June), Fall (July-December)

**Synthesis Theme**:

- [X] Multiple System Drivers


```{r autostats_New Indicator}
# Either from Contributor or ecodata
```

## Implications
Distribution shifts for zooplankton may result in mismatches with zooplankton feeding fish, marine mammal, and seabirds.

## Get the data

**Point of contact**: [Sarah Gaichas Sarah.Gaichas@noaa.gov](mailto:Sarah Gaichas Sarah.Gaichas@noaa.gov){.email}

**ecodata name**: No dataset

**Variable definitions**

Waiting until we aggregate the datasets, if formatted similarly to forage and benthos, the cog data will be part of the zooplankton indices dataset


No Data

**Indicator Category**:

- [X] Extensive analysis, not yet published


## Public Availability

Source data are NOT publicly available.

## Accessibility and Constraints

Request from Harvey Walsh, harvey.walsh@noaa.gov

