# Copepod Preyfield Energy Density {#preyfield_energy}

```{r echo=FALSE}
knitr::opts_chunk$set(echo = F)
library(ecodata)
library(downloadthis)
```

*Last updated March 30, 2026*

**Description**: Coming soon

**Contributor(s)**: Coming soon

**Affiliations**: NEFSC

**Indicator Family**: 




**Indicator Category**:




**Synthesis Theme**:




```{r download_preyfield_energy}
# Download dataset from ecodata
try(downloadthis::download_this(
ecodata::preyfield_energy,
output_name = 'preyfield_energy',
output_extension = '.csv',
button_label = 'Download dataset from `ecodata`',
button_type = 'default',
has_icon = TRUE,
icon = 'fa fa-save',
class = 'hvr-sweep-to-left',
csv2 = F),
silent = TRUE)
```

## Introduction to Indicator
Coming soon

## Key Results and Visualizations
Coming soon

## Implications
Coming soon

## Indicator statistics 
Spatial scale: Coming soon

Temporal scale: Coming soon

**Variable definitions**

```{r vars_preyfield_energy}
# Pull all var names
vars <- ecodata::preyfield_energy |>
   dplyr::select(Var) |>
   dplyr::distinct()

DT::datatable(vars)
```
## Get the data

**Point of contact**: [EDAB (edab.data@noaa.gov)](mailto:EDAB (edab.data@noaa.gov)){.email}

**ecodata dataset**: `ecodata::preyfield_energy`

**Tech Doc link**: <https://noaa-edab.github.io/tech-doc/preyfield_energy.html>

### Public Availability

Source data are publicly available.

