# Copepod Preyfield Energy Density {#preyfield_energy}

**Description**: Coming soon

**Indicator family**: 




**Contributor(s)**: Coming soon

**Affiliations**: NEFSC

```{r echo=FALSE}
knitr::opts_chunk$set(echo = F)
library(ecodata)
```
## Introduction to Indicator
Coming soon

## Key Results and Visualizations
Coming soon


## Indicator statistics 
Spatial scale: Coming soon

Temporal scale: Coming soon

**Synthesis Theme**:




```{r autostats_preyfield_energy}
# Either from Contributor or ecodata
```

## Implications
Coming soon

## Get the data

**Point of contact**: [EDAB (edab.data@noaa.gov)](mailto:EDAB (edab.data@noaa.gov)){.email}

**ecodata name**: `ecodata::preyfield_energy`

**Variable definitions**

Coming soon

```{r vars_preyfield_energy}
# Pull all var names
vars <- ecodata::preyfield_energy |>
   dplyr::select(Var) |>
   dplyr::distinct()

DT::datatable(vars)
```
**Indicator Category**:




## Public Availability

Source data are publicly available.

## Accessibility and Constraints

_No response_

**tech-doc link**
<https://noaa-edab.github.io/tech-doc/preyfield_energy.html>

