# Species Distribution Indicators {#species_dist}

**Description**: Species mean depth, along-shelf distance, and distance to coastline

**Indicator family**: 

- [X] Habitat
- [X] Megafauna


**Contributor(s)**: Kevin Friedland, Brandon Beltz

**Affiliations**: NEFSC

```{r echo=FALSE}
knitr::opts_chunk$set(echo = F)
library(ecodata)
```
## Introduction to Indicator
Distribution shifts for a suite of 48 commercially or ecologically important fish species were evaluated using center of gravity metrics based on NEFSC bottom trawl survey data.

Along-shelf distance is a metric for quantifying the distribution of a species through time along the axis of the US Northeast Continental Shelf, which extends northeastward from the Outer Banks of North Carolina. Once mean distance is found, depth of occurrence and distance to coastline can be calculated for each species’ positional center.

Published distribution metrics for highly migratory species are also available for comparison.

## Key Results and Visualizations
Species distribution shifts have been noted for several highly migratory species, including sharks, billfish and tunas between 2002 and 2019 (Fig. 1) [@crear_climate-influenced_2023].

![Figure 1. The median location of catch each year for albacore tuna (a), bigeye tuna (b), large bluefin tuna (c), small bluefin tuna (d), skipjack tuna (e), yellowfin tuna (f), blue shark (g), shortfin mako shark (h), thresher shark (i), blue marlin (j), swordfish (k), and white marlin (l). The color of the dot indicates the year, and the error bars represent the upper and lower quartiles for a given year. To reduce overlap among median catch locations, a small jitter was added to the median latitude and longitude when appropriate. Source: reprinted from Crear et al. 2023 (Fig 3)](https://github.com/NOAA-EDAB/ecodata/raw/dev/data-raw/workshop/images/LPSFig_Spatial_HighRes-JenCudney_2025.png){width=100%}


The center of distribution for a suite of 48 commercially or ecologically important fish species along the entire Northeast Shelf continues to show movement towards the northeast and generally into deeper water.

### MidAtlantic

```{r plot_species_distMidAtlanticalong}
# Plot indicator
ggplotObject <- ecodata::plot_species_dist(report= 'MidAtlantic', varName= 'along',n=10)
ggplotObject
```

```{r plot_species_distMidAtlanticdepth}
# Plot indicator
ggplotObject <- ecodata::plot_species_dist(report= 'MidAtlantic', varName= 'depth',n=10)
ggplotObject
```

### NewEngland

```{r plot_species_distNewEnglandalong}
# Plot indicator
ggplotObject <- ecodata::plot_species_dist(report= 'NewEngland', varName= 'along',n=10)
ggplotObject
```

```{r plot_species_distNewEnglanddepth}
# Plot indicator
ggplotObject <- ecodata::plot_species_dist(report= 'NewEngland', varName= 'depth',n=10)
ggplotObject
```


## Indicator statistics 
Spatial scale: Shelfwide

Temporal scale: Annual

**Synthesis Theme**:

- [X] Multiple System Drivers


```{r autostats_species_dist}
# Either from Contributor or ecodata
```

## Implications
Temperature change is a major driver of changing fish distributions [@friedland_event_2019].

## Get the data

**Point of contact**: [Kevin Friedland, kevin.friedland@noaa.gov](mailto:Kevin Friedland, kevin.friedland@noaa.gov){.email}

**ecodata name**: `ecodata::species_dist`

**Variable definitions**

"along-shelf distance"  "depth"                 "distance to coast"     "Latitude"              "Longitude"

```{r vars_species_dist}
# Pull all var names
vars <- ecodata::species_dist |>
   dplyr::select(Var, Units) |>
   dplyr::distinct()

DT::datatable(vars)
```
**Indicator Category**:

- [X] Published Methods


## Public Availability

Source data are NOT publicly available.

## Accessibility and Constraints

Contact Kevin Friedland, kevin.friedland@noaa.gov for data access

**tech-doc link**
<https://noaa-edab.github.io/tech-doc/species_dist.html>

