# Bottom trawl survey diversity index {#survey_shannon}

**Description**: Diversity metric for the bottom trawl survey data

**Indicator family**: 

- [X] Megafauna


**Contributor(s)**: Andy Beet, Sean Lucey

**Affiliations**: NEFSC

```{r echo=FALSE}
knitr::opts_chunk$set(echo = F)
library(ecodata)
```
## Introduction to Indicator
Proceed as though this were a short summary of a typical introduction section in a paper.

## Key Results and Visualizations
Proceed as though this were a short summary of a typical results section in a paper.

### MAB

```{r plot_survey_shannonMAB}
# Plot indicator
ggplotObject <- ecodata::plot_survey_shannon(report='MidAtlantic',n=10)
ggplotObject
```

### NE

```{r plot_survey_shannonNE}
# Plot indicator
ggplotObject <- ecodata::plot_survey_shannon(report='NewEngland',n=10)
ggplotObject
```


## Indicator statistics 
Spatial scale: EPU

Temporal scale: SPRING, FALL

**Synthesis Theme**:

- [X] Regime Shifts
- [X] Ecosystem Reorganization


```{r autostats_survey_shannon}
# Either from Contributor or ecodata
```

## Implications
Proceed as though this were a short summary of a typical discussion section in a paper.

## Get the data

**Point of contact**: [andrew.beet@noaa.gov](mailto:andrew.beet@noaa.gov){.email}

**ecodata name**: `ecodata::survey_shannon`

**Variable definitions**

Name: NEFSC survey species diversity (SPRING, FALL) 
Units: Shannon index (unitless)

```{r vars_survey_shannon}
# Pull all var names
vars <- ecodata::survey_shannon |>
   dplyr::select(Var, Units) |>
   dplyr::distinct()

DT::datatable(vars)
```
**Indicator Category**:

- [X] Published Methods
- [X] Database pull


## Public Availability

Source data are publicly available.

## Accessibility and Constraints

_No response_

**tech-doc link**
<https://noaa-edab.github.io/tech-doc/survey_shannon.html>

