# Thermal Habitat Persistence {#thermal_habitat_gridded}

```{r echo=FALSE}
knitr::opts_chunk$set(echo = F)
library(ecodata)
library(downloadthis)
```

*Last updated April 03, 2026*

**Description**: The number of days per year per 1/12 degree cell that exceeds a temperature threshold.

**Contributor(s)**: Joe Caracappa

**Affiliations**: NEFSC

**Indicator Family**: 

- [X] Oceanographic
- [X] Habitat


**Indicator Category**:

- [X] Extensive analysis, not yet published


**Synthesis Theme**:

- [X] Multiple System Drivers


```{r download_thermal_habitat_gridded}
# Download dataset from ecodata
try(downloadthis::download_this(
ecodata::thermal_habitat_gridded,
output_name = 'thermal_habitat_gridded',
output_extension = '.csv',
button_label = 'Download dataset from `ecodata`',
button_type = 'default',
has_icon = TRUE,
icon = 'fa fa-save',
class = 'hvr-sweep-to-left',
csv2 = F),
silent = TRUE)
```

## Introduction to Indicator
Many deep water benthic and demersal species exhibit thermal preferences for metabolic, reproductive, and growth processes. Temperatures above these thermal presences may impair these processes. Data originate from GLORYS12V1 global reanalysis for 1993-01-01 to 2023-08-29, and from PSY forecasts from 2021-01-01 to 2023-12-31. Cells are mapped to 3 depth bins: 0-25m, 25-100m, and >100m. Two temperature thresholds are used, representing a temperature where moderate (15C) and extreme (24C) thermal stresses are likely to occur across several species. GLORYS and PSY 1/12 degree grid is mapped to EPU_NOESTUARIES shape files by the center point of each grid cell.

## Key Results and Visualizations
Maps of full NE shelf with cell shading gradient showing number of days exceeding temperature thresholds. For 15C threshold, much of the southern Mid-Atlantic Bight (MAB), the shelf break and Nantucket Shoals experienced >100 days. In 2023, only the southern and/or shallow portion of the MAB exceeded the 24C threshold.

**Mid-Atlantic**
```{r plot_thermal_habitat_griddedMidAtlantic}
# Plot indicator
ggplotObject <- ecodata::plot_thermal_habitat_gridded(report = 'MidAtlantic')
ggplotObject
try(downloadthis::download_this(
ggplotObject$data,
output_name = 'thermal_habitat_griddedMidAtlantic',
output_extension = '.csv',
button_label = 'Download the data plotted above',
button_type = 'default',
has_icon = TRUE,
icon = 'fa fa-save',
class = 'hvr-sweep-to-left',
csv2 = F),
silent = TRUE)
```
<br>

**New England**
```{r plot_thermal_habitat_griddedNewEngland}
# Plot indicator
ggplotObject <- ecodata::plot_thermal_habitat_gridded(report = 'NewEngland')
ggplotObject
try(downloadthis::download_this(
ggplotObject$data,
output_name = 'thermal_habitat_griddedNewEngland',
output_extension = '.csv',
button_label = 'Download the data plotted above',
button_type = 'default',
has_icon = TRUE,
icon = 'fa fa-save',
class = 'hvr-sweep-to-left',
csv2 = F),
silent = TRUE)
```
<br>

## Implications
Using a high resolution bottom temperature product elucidates localized heat stress within EPUs that may be masked when looking at mean EPU bottom temperature. This can be important to sessile species who aren't able to move out of warm water conditions.

## Indicator statistics 
Spatial scale: full shelf

Temporal scale: annual

**Variable definitions**

```{r vars_thermal_habitat_gridded}
# Pull all var names
vars <- ecodata::thermal_habitat_gridded |>
   dplyr::select(Var, Units) |>
   dplyr::distinct()

DT::datatable(vars)
```
## Get the data

**Point of contact**: [joseph.caracappa@noaa.gov](mailto:joseph.caracappa@noaa.gov){.email}

**ecodata dataset**: `ecodata::thermal_habitat_gridded`

**Tech Doc link**: <https://noaa-edab.github.io/tech-doc/thermal_habitat_gridded.html>

### Public Availability

Source data are publicly available.

