# Thermal Habitat Area {#thermal_habitat_area}

```{r echo=FALSE}
knitr::opts_chunk$set(echo = F)
library(ecodata)
library(downloadthis)
```

*Last updated February 20, 2025*

**Description**: Calculates the proportion of each EPU that exceeds temperature thresholds as a daily time series from 1993 – 2024

**Contributor(s)**: Joe Caracappa

**Affiliations**: NEFSC

**Indicator Family**: 

- [X] Oceanographic
- [X] Habitat


**Indicator Category**:

- [X] Extensive analysis, not yet published


**Synthesis Theme**:

- [X] Multiple System Drivers


```{r download_thermal_habitat_area}
# Download dataset from ecodata
downloadthis::download_this(
ecodata::thermal_habitat_area,
output_name = 'thermal_habitat_area',
output_extension = '.csv',
button_label = 'Download dataset from `ecodata`',
button_type = 'default',
has_icon = TRUE,
icon = 'fa fa-save',
class = 'hvr-sweep-to-left',
csv2 = F)
```

## Introduction to Indicator
See "Thermal Habitat Persistence" for cell-based calculations. Many deep water benthic and demersal species exhibit thermal preferences for metabolic, reproductive, and growth processes. Temperatures above these thermal preferences may impair these processes. Two temperature thresholds (15 and 24 degrees C) were chosen based on cutoff points where demersal species are less commonly found. Thermal habitat area is calculated by identifying 1/12 degree cells within a given EPU that are greater than or equal to the temperature threshold then taking the sum of all cell areas. Data from 1993-01-01 to 2024-09-10 were obtained from the CMEMS' GLORYS12V1 global reanalysis product.

## Key Results and Visualizations
For each EPU, bottom temperature data were separated by depth bin (0-25m; 25m-100m; and >100m), as well as temperature threshold (>15C , >24C), and the proportion of each EPU exceeding these thresholds was calculated on a daily basis from 1993 to 2024. Time series of each year's proportional area show an increasing proportion of the MAB and GB mid-depths have experienced high temperatures in recent years. The 24C threshold is rarely seen in GB and GOM.

**Mid-Atlantic Bight**
```{r plot_thermal_habitat_areaMABMidAtlantic}
# Plot indicator
ggplotObject <- ecodata::plot_thermal_habitat_area(EPU = 'MAB', report = 'MidAtlantic')
ggplotObject
try(downloadthis::download_this(
ggplotObject$data,
output_name = 'thermal_habitat_areaMABMidAtlantic',
output_extension = '.csv',
button_label = 'Download the data plotted above',
button_type = 'default',
has_icon = TRUE,
icon = 'fa fa-save',
class = 'hvr-sweep-to-left',
csv2 = F),
silent = TRUE)
```
<br>

**Georges Bank**
```{r plot_thermal_habitat_areaGBNewEngland}
# Plot indicator
ggplotObject <- ecodata::plot_thermal_habitat_area(EPU = 'GB', report = 'NewEngland')
ggplotObject
try(downloadthis::download_this(
ggplotObject$data,
output_name = 'thermal_habitat_areaGBNewEngland',
output_extension = '.csv',
button_label = 'Download the data plotted above',
button_type = 'default',
has_icon = TRUE,
icon = 'fa fa-save',
class = 'hvr-sweep-to-left',
csv2 = F),
silent = TRUE)
```
<br>

**Gulf of Maine**
```{r plot_thermal_habitat_areaGOMNewEngland}
# Plot indicator
ggplotObject <- ecodata::plot_thermal_habitat_area(EPU = 'GOM', report = 'NewEngland')
ggplotObject
try(downloadthis::download_this(
ggplotObject$data,
output_name = 'thermal_habitat_areaGOMNewEngland',
output_extension = '.csv',
button_label = 'Download the data plotted above',
button_type = 'default',
has_icon = TRUE,
icon = 'fa fa-save',
class = 'hvr-sweep-to-left',
csv2 = F),
silent = TRUE)
```
<br>

## Implications
If a large proportion of species current habitat become thermally inhospitable, it can influence species' productivity and in, extreme cases, cause mortality.

## Indicator statistics 
Spatial scale: EPU

Temporal scale: daily

**Variable definitions**

```{r vars_thermal_habitat_area}
# Pull all var names
vars <- ecodata::thermal_habitat_area |>
   dplyr::select(Var, Units) |>
   dplyr::distinct()

DT::datatable(vars)
```
## Get the data

**Point of contact**: [joseph.caracappa@noaa.gov](mailto:joseph.caracappa@noaa.gov){.email}

**ecodata dataset**: `ecodata::thermal_habitat_area`

**Tech Doc link**: <https://noaa-edab.github.io/tech-doc/thermal_habitat_area.html>

### Public Availability

Source data are publicly available.

